Relative Stable Maps in Genus One via Central Alignments

### 2019-01-30 GenusOneRelativeMaps.tex
minor changes

\begin{abstract} For a smooth projective variety $X$ and a smooth very ample hypersurface $Y \subseteq X$, we define moduli spaces of relative stable maps to $(X,Y)$ in genus one, as closed substacks of the moduli space of maps from centrally aligned curves, constructed in \cite{RSPW}. We construct virtual classes for these moduli spaces, which we use to define \emph{reduced relative Gromov--Witten invariants} in genus one.

@@

\subsection{Blowing up the Artin fan}
Let $\Mcal = \ol{\Mcal}^{\op{log}}_{1,\alpha}(\PP^N|H,d)$ denote the Abramovich--Chen--Gross--Siebert moduli space of log stable maps. Recall that $\VZ$ is obtained as a closed substack of a log modification
\begin{equation*} \widetilde{\VZ} \to \Mcal .\end{equation*}
Since the map $\VZ \hookrightarrow \widetilde{\VZ}$ is strict, the Artin fan $\Acal$ of $\VZ$ is locally isomorphic to the Artin fan of $\widetilde{\VZ}$ (with the latter being, in general, larger). Since the following discussion is entirely local, we will ignore the distinction between the two, and pretend as though $\Acal$ is the Artin fan of $\widetilde{\VZ}$. There is then a commuting square:
\bcd
\widetilde{\VZ} \ar[r] \ar[d] & \Mcal \ar[d] \\
\Acal \ar[r] & \Acal_{\Mcal}.
\ecd
Note that neither of the vertical maps are smooth, since the moduli spaces on the top row are not log smooth. The construction of $\widetilde{\VZ} \to \Mcal$ as the log modification obtained by imposing an alignment condition gives us a combinatorial description of the map $\Acal \to \Acal_{\Mcal}$ which we will use to study $\Acal$. Recall that $\Acal_\Mcal$ is locally isomorphic to the stack quotient
\begin{equation*} \left[ \Spec \kfield[Q] / \Spec\kfield [Q^{\gp}] \right]\end{equation*}
where $Q$ is a monoid giving a\footnote{(Navid) Neat?} local chart for $\Mcal$, which we may take to be the minimal monoid of \cite[\S 1.5]{GrossSiebertLog}. The real dual $Q^\vee_{\RR} = \Hom(Q,\RR_{\geq 0})$ of this monoid can be viewed as a moduli space of tropical maps; see \cite[Remark 1.12]{GrossSiebertLog}. We call this \emph{the tropical moduli space}; contained in the tropical moduli space are the edge lengths of the associated tropical curve (corresponding to the smoothing parameters of the nodes of the logarithmic curve). Since the alignment condition amounts to imposing a partial ordering amongst certain sums of these edge lengths, it produces a polyhedral decomposition of the tropical moduli space, into chambers where different partial order relations hold. If we only consider the integral points, this produces a polyhedral decomposition of the cone $Q^\vee = \Hom(Q,\N)$. Dualising, we obtain a toric blow-up
\begin{equation*} Z \to \Spec \kfield[Q] \end{equation*}
which, since it is equivariant, descends to a morphism of the associated zero-dimensional stacks:
\begin{equation*} \left[Z/T_Z \right] \to \left[ \Spec \kfield[Q] / \Spec\kfield [Q^{\gp}] \right]. \end{equation*}
This gives a local description of the map $\Acal \to \Acal_\Mcal$ and in particular of the Artin fan $\Acal$. By the orbit-cone correspondence, the codimension-$k$ strata of $\Acal$ which intersect this open locus correspond to the $k$-dimensional cones in the polyhedral decomposition of $Q^\vee$ described above. These can be understood entirely in terms of tropical combinatorics. This is best explained through a number of examples, which we now present.

@@

\subsection{Splitting and gluing punctured logarithmic maps}\label{subsection punctured maps} We still start by considering the usual moduli space $\Mcal$ of log stable maps. In \cite{PuncturedMaps}, a recursive description of the logarithmic strata of $\Mcal$ is given, in terms of moduli spaces of punctured maps. As discussed above, the logarithmic strata $\Scal$ of $\Mcal$ are indexed by combinatorial types of tropical curves. Given such a combinatorial type $\Delta$, each vertex $v$ of the associated graph defines a moduli space of punctured stable maps:
\begin{equation*} \ol\Mcal_v = \Mpunct{g(v)}{\alpha(v)}{Z_v}{d_v}. \end{equation*}
Here $g(v)$ and $d(v)$ are the genus and degree assigned to the vertex, while $Z_v$ is the logarithmic stratum of the target corresponding to the tropical stratum into which the vertex is mapped by the combinatorial type (in our setting, $Z_v=\PP^N$ or $H$). The length of the vector $\alpha(v)$ is given by the number of edges and flags adjacent to $v$, and the entries (which represent tangencies, and may be positive or negative) are given by the expansion factors along those edges and flags. Note that if all of the tangencies are positive, then this is nothing but an ordinary moduli space of log stable maps. The corresponding logarithmic stratum is then obtained as a fibre product over evaluation maps:
\bcd
\Scal \ar[r] \ar[d] \ar[rd,phantom,"\square"] & \prod_v \ol\Mcal_v \ar[d] \\
\prod_{e} Z_e \ar[r] & \prod_{e} Z_e\times Z_e.
\ecd
This fibre product must be take in the fs log category. However, in our setting, the ordinary fibre product is automatically integral\footnote{(Navid) Justify this}, so all that is required is a saturation. This means that the fs fibre product is just a finite cover of the ordinary fibre product, and we may calculate the degree directly.\footnote{(Navid) How, again?}

>>>>

\begin{abstract} [OLD] For a smooth projective variety $X$ and a smooth very ample hypersurface $Y \subseteq X$, we define moduli spaces of relative stable maps to $(X,Y)$ in genus one, as closed substacks of the moduli space of maps from centrally aligned curves, constructed in \cite{RSPW}. We construct virtual classes for these moduli spaces, which we use to define \emph{reduced relative Gromov--Witten invariants} in genus one.

@@

\subsection{Blowing up the Artin fan}
Let $\Mcal = \ol{\Mcal}^{\op{log}}_{1,\alpha}(\PP^N|H,d)$ denote the Abramovich--Chen--Gross--Siebert moduli space of log stable maps. Recall that $\VZ$ is obtained as a closed substack of a log modification
\begin{equation*} \widetilde{\VZ} \to \Mcal .\end{equation*}
Since the map $\VZ \hookrightarrow \widetilde{\VZ}$ is strict, the Artin fan $\Acal$ of $\VZ$ is locally isomorphic to the Artin fan of $\widetilde{\VZ}$ (with the latter being, in general, larger)\footnote{(Navid) Make this more precise}. Since the following discussion is entirely local, we will ignore the distinction between the two, and pretend as though $\Acal$ is the Artin fan of $\widetilde{\VZ}$. There is then a commuting square:
\bcd
\widetilde{\VZ} \ar[r] \ar[d] & \Mcal \ar[d] \\
\Acal \ar[r] & \Acal_{\Mcal}.
\ecd
Note that neither of the vertical maps are smooth, since the moduli spaces on the top row are not log smooth. The construction of $\widetilde{\VZ} \to \Mcal$ as the log modification obtained by imposing an alignment condition gives us a combinatorial description of the map $\Acal \to \Acal_{\Mcal}$ which we will use to study $\Acal$. Recall that $\Acal_\Mcal$ is locally isomorphic to the stack quotient
\begin{equation*} \left[ \Spec \kfield[Q] / \Spec\kfield [Q^{\gp}] \right]\end{equation*}
where $Q$ is a monoid giving a\footnote{(Navid) Neat?} local chart for $\Mcal$, which we may take to be the minimal monoid of \cite[\S 1.5]{GrossSiebertLog}. The real dual $Q^\vee_{\RR} = \Hom(Q,\RR_{\geq 0})$ of this monoid can be viewed as a moduli space of tropical maps; see \cite[Remark 1.12]{GrossSiebertLog}. We call this \emph{the tropical moduli space}; contained in the tropical moduli space are the edge lengths of the associated tropical curve (corresponding to the smoothing parameters of the nodes of the logarithmic curve). Since the alignment condition amounts to imposing a partial ordering amongst certain sums of these edge lengths, it produces a polyhedral decomposition of the tropical moduli space, into chambers where different partial order relations hold. If we only consider the integral points, this produces a polyhedral decomposition of the cone $Q^\vee = \Hom(Q,\N)$. Dualising, we obtain a toric blow-up
\begin{equation*} Z \to \Spec \kfield[Q] \end{equation*}
which, since it is equivariant, descends to a morphism of the associated zero-dimensional stacks:
\begin{equation*} \left[Z/T_Z \right] \to \left[ \Spec \kfield[Q] / \Spec\kfield [Q^{\gp}] \right]. \end{equation*}
This gives a local description of the map $\Acal \to \Acal_\Mcal$ and in particular of the Artin fan $\Acal$. By the orbit-cone correspondence, the codimension-$k$ strata of $\Acal$ which intersect this open locus correspond to the $k$-dimensional cones in the polyhedral decomposition of $Q^\vee$ described above. These can be understood entirely in terms of tropical combinatorics. This is best explained through a number of examples, which we now present.

@@

\subsection{Splitting and gluing punctured logarithmic maps}\label{subsection punctured maps} We still start by considering the usual moduli space $\Mcal$ of log stable maps. In \cite{PuncturedMaps}, a recursive description of the logarithmic strata of $\Mcal$ is given, in terms of moduli spaces of punctured maps. As discussed above, the logarithmic strata $\Scal$ of $\Mcal$ are indexed by combinatorial types of tropical curves. Given such a combinatorial type $\Delta$, each vertex $v$ of the associated graph defines a moduli space of punctured stable maps:
\begin{equation*} \ol\Mcal_v = \Mpunct{g(v)}{\alpha(v)}{Z_v}{d_v}. \end{equation*}
Here $g(v)$ and $d(v)$ are the genus and degree assigned to the vertex, while $Z_v$ is the logarithmic stratum of the target corresponding to the tropical stratum into which the vertex is mapped by the combinatorial type (in our setting, $Z_v=\PP^N$ or $H$). The length of the vector $\alpha(v)$ is given by the number of edges and legs adjacent to $v$, and the entries (which represent tangencies, and may be positive or negative) are given by the expansion factors along those edges and legs. Note that if all of the tangencies are positive, then this is nothing but an ordinary moduli space of log stable maps. The corresponding logarithmic stratum is then obtained as a fibre product over evaluation maps:
\bcd
\Scal \ar[r] \ar[d] \ar[rd,phantom,"\square"] & \prod_v \ol\Mcal_v \ar[d] \\
\prod_{e} Z_e \ar[r] & \prod_{e} Z_e\times Z_e.
\ecd
This fibre product must be take in the fs log category. However, in our setting, the ordinary fibre product is automatically integral\footnote{(Navid) Justify this}, so all that is required is a saturation. This means that the fs fibre product is just a finite cover of the ordinary fibre product, and we may calculate the degree directly.\footnote{(Navid) How, again?}
